---
title: "Data 607 Project 1"
author: "Mikhail bBroomes"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

In this project, you’re given a text file with chess tournament results where the information has some structure. Your
job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database)
with the following information for all of the players:

## Libraries
```{r setup, include=FALSE}
library(tidyverse)
library(readr)
```

## Loading Data 


```{r cars}
chess_data <- readLines("https://raw.githubusercontent.com/MAB592/DATA-607-Projects-/main/tournamentinfo.txt")
print(chess_data)
```

## String Extraction 

Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents

```{r pressure, echo=FALSE}
player_name <- unlist(str_extract_all(chess_data,"(?<=\\|)[^|]*(?=\\s\\|\\d)"))
player_state <- unlist(str_extract_all(chess_data,"(?<=\\s{3})[:upper:]{2}(?=\\s*\\|\\s)"))
player_num <-  as.numeric(unlist(str_extract_all(chess_data,"(?<=\\s{3,4})\\d{1,2}(?=\\s)")))
player_totalpts <- as.numeric(unlist(str_extract_all(chess_data, "(?<=\\|)\\d\\.\\d")))
player_pre_rating <- as.numeric(unlist(str_extract_all(chess_data, "(?<=R:\\s{1,2})(\\d{3,4}(?=\\s))|(\\d{3,4}(?=P\\d{1,2}\\s*-))")))
print (player_name)
```
Creating the data frame from the extracted chess data

```{r}
processed_data <- data.frame(player_num, player_name, player_state, player_totalpts, player_pre_rating)

str(processed_data)
```



```{r} 
opp_avg = as.numeric(unlist(str_extract_all(chess_data,"(?<=W|D|H)\\s{2,3}(\\d{1,2})(?=\\|)")))
print (opp_avg)
```
## Finding Opponent Average 


Extracting for the averages 


```{r}
secondary_rows <- chess_data[seq(5, 196, 3)]
opponent_num <- as.numeric(unlist(str_extract_all(secondary_rows, "(?<=\\|(W|L|D)\\s{2,3})[[:digit:]]{1,2}(?=\\|)|((?<!->)(?<=\\|(U|H|B|X))\\s{4}(?=\\|))")))
print(secondary_rows)
```

## Calculating Opponent Average


```{r}
# Create matrix to store data calculated in the for loop.  Pre-populating values with NA for more efficient processing in R.
pcr_matrix <- matrix(data = NA, nrow = 64, ncol = 2)

colnames(pcr_matrix) <- c("total_opp_pcr", "avg_opp_pcr")

row_counter <- 0

for(i in seq(from=1, to=length(opponent_num)-6, by=7)){
  row_counter <- row_counter + 1
  
# Perform a lookup of each competitor's score based on their player number and add the up for each row (corresponding to each sequence of 7 data points, w/ value from for loop serving as row 'anchor')
  pcr_matrix[row_counter, 1] <- (sum(subset(processed_data$player_pre_rating, processed_data$player_num %in% opponent_num[seq(from=i, to=i+6, by=1)])))
  
# Calculate the average score for each row, excluding missing entries
  pcr_matrix[row_counter, 2] <- pcr_matrix[row_counter, 1] / length(subset(opponent_num[seq(from=i, to=i+6, by=1)],!is.na(opponent_num[seq(from=i, to=i+6, by=1)])))
  
}
# Check matrix was processed properly by looking at the first few rows of output
head(pcr_matrix, 5)
```
```{r}
# Round the figures to the nearest whole number
pcr_matrix[, 2] <- round(pcr_matrix[,2], digits = 0)

# Add average scores to data frame with other processed data and rename for readability
processed_data <- cbind(processed_data, pcr_matrix[, 2])
processed_data <- rename(processed_data, avg_opp_pcr = `pcr_matrix[, 2]`)
```

path <- getwd()

write.csv(processed_data, file.path(path, "chess_processed_data.csv"))

```{r}
excelfile <- read_csv("C:\\Users\\mikha\\OneDrive\\Desktop\\Data 607\\chess_processed_data.csv")
print(excelfile)
```


path <- getwd()

write.csv(processed_data, file.path(path, "chess_processed_data.csv"))

https://rpubs.com/MikhailB/1088011