---
title: "Data 607 Assignment 3"
author: "Mikhail Broomes"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
## Libraries 

```{r}
library (tidyverse) 

```

```{r setup, include=TRUE}
majors <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/all-ages.csv")

```


## Question 1 

Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r cars}

majors_with_data_or_stats <- majors %>%
  filter(str_detect(Major, regex("DATA|STATISTICS", ignore_case = TRUE)))

print(majors_with_data_or_stats)

```

## Question 2 

Write code that transforms the data below:
[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"
Into a format like this:
c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")


```{r pressure, echo=TRUE}

clean_data <- c("bell pepper", "bilberry", "blackberry", "blood orange", 
                "blueberry", "cantaloupe", "chili pepper", "cloudberry", 
                "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

untransformed_data <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
  
  [5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"
  
  [9] "elderberry"   "lime"         "lychee"       "mulberry" 
  
  [13] "olive"        "salal berry"'

transformed_data <- unlist(str_extract_all(untransformed_data, '[[:alpha:]]+\\s[[:alpha:]]+|[[:alpha:]]+'))


transformed_data

```
## Question 3 

Describe, in words, what these expressions will match:

(..)\1

This references two characters (.)followed by the same characters. An extra backslash was need as to being interpreted as a literal backslash

```{r}
text3 <- "aedaddaxyzabab"
str_view(text3, "(..)\\1", match = TRUE)
```

"(.).\\1.\\1"


The reference is looking for a sequence where there is a character, followed by any character, followed by any character, followed by the same character as the first one, and then followed by any character, with the first character repeating once more in the sequence.
```{r}
text4 <- "dcbabazadcbsdsms"
str_view(text4, "(.).\\1.\\1", match = TRUE)
```

"(.)(.)(.).*\\3\\2\\1"

The reference matches a sequence in a string where the first three characters are followed by any characters and then the same three characters appear in reverse order

```{r}
text5 <- "stuwyqabczcbaioeuo"
str_view(text5, "(.)(.)(.).*\\3\\2\\1", match = TRUE)
```

## Question 4

Construct regular expressions to match words that:

Start and end with the same character.

```{r}
words <- c("banana", "apple", "cherry", "radar", "level", "pop")

str_detect(words,"^(.).*\\1$")
```

Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

```{r}
letters <- c("church", "success", "committee", "occurred", "aggressive", "necessary")
str_detect(letters,"(..).*\\1")
```

Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

```{r}
letters2 <- c("church", "success", "committee", "occurred", "aggressive", "necessary", "eleven","accelerate" )
str_detect(letters2,"(.).*\\1.*\\1.*")
```

