---
title: "Data 607 assignment 5"
author: "Mikhail bBroomes"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

## Libraries 
```{r setup, include=TRUE}

library (tidyverse)
library(readr)

```

## Sample data 

Using Github to extra my data to be analyze

```{r cars}
data <- read.csv("https://raw.githubusercontent.com/MAB592/Data-607-Assignments/main/Airline_wk5%20-%20Sheet1.csv")
print (data)
```

## Renaming Columns 

Here I am renaming the columns where I see fit

```{r pressure, echo=TRUE}
rename_data <-  data %>% 
  rename("Flight Status" = Flight.Status,
         "Los Angeles" = Los.Angeles,
         "San Diego" = San.Diego,
         "San Francisco" = San.Francisco
         
  )
  
print(rename_data) 
```

I am now converting my data to a more longer format to make it easier to graph my data 

```{r}
long_data <- rename_data %>% 
  pivot_longer (
    cols = c("Los Angeles", "Phoenix", "San Diego", "San Francisco", "Seattle"),
    names_to = "Location",
    values_to = "Amount of flights"
  )

print(long_data)
```


Filling the missing data from the Airline column that contains NA using the mutate function 


```{r}
final_data <- long_data %>% 
  mutate(Airline = as.character(na_if(Airline, 'NA'))) %>%  
  fill(Airline, .direction = 'down')

print(final_data)
```


Now filtering for delayed flights. There was an error when I was just using the filter function where it returned 0, so I used the function trimws in order to produce my required table.

```{r}
final_data$`Flight Status` <- trimws(final_data$`Flight Status`)

delayed_data <- final_data %>%
  filter(`Flight Status` == "delayed")

print(delayed_data)
```
## Analysis

Using ggplot to look at the data for our delayed flights we see that AM West has the larger amount of delayed flights 

```{r}
g <- ggplot(delayed_data,aes(x = `Airline`,y = `Amount of flights`,fill = Airline ))+ 
  geom_bar(stat='identity' ) + 
  labs(
    title = "Airlines with Delayed Flights",
    x = "Airline",
    y = "Amount of Delayed Flights"
  )

print (g) 
```
Now in order to gain some insights using the statistical packages in R I made the delayed text have its own column in order to look at the different locations and see location has the most delays. As we can see Phoenix has the highest average delays among both airlines.

```{r}
delayed_statistics <- delayed_data %>%
  pivot_wider(names_from = `Flight Status`, values_from = `Amount of flights`)

print(delayed_statistics)


delayed_statistics  %>% 
        group_by(Location) %>% 
        dplyr::summarise(max = max(delayed),
        min=min(delayed),
        mean=mean(delayed),  
        median=median(delayed)
        )
```


